Transfer of Learned Opponent Models in Zero Sum Games

### 2021-06-04 Report/CSS_submission_2021.Rmd
added report in papaja and softmax to BCH models

---
title             : "Transfer of Learned Opponent Models in Zero Sum Games"
shorttitle        : "Draft-3"

@@

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
linkcolor         : "blue"
mask              : no
draft             : no
fig_caption       : yes

documentclass     : "apa6"
# classoption       : a4paper
classoption       : man
output            : papaja::apa6_pdf
#citation_package  : natbib
---

@@

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(plyr)
library(tidyr)
library(dplyr)
library(MASS)
library(ggpubr)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(magick)
library(depmixS4)
library(rmdfiltr)
```

@@

```{r, include=FALSE} 
options(tinytex.verbose = TRUE)
add_wordcount_filter(NULL, error = FALSE)
```

# Introduction
Being able to transfer previously acquired knowledge to a new domain is one of the hallmarks of human intelligence. Humans are naturally endowed with the ability to extract relevant features from a situation, identify the presence of these features in a novel setting and use previously acquired knowledge to adapt to previously unseen challenges using acquired knowledge. More formally, @perkins1992transfer defines transfer of learning as the application of skills, knowledge, and/or attitudes that were learned in one situation to another learning situation. This typically human skill has so far eluded modern AI agents. Deep neural networks for instance can do very well on image recognition tasks and can even reach super-human performance levels on video and strategic board games. Yet they struggle to learn as fast or as efficiently as humans do, and more importantly they have a very limited ability to generalize and transfer knowledge to new domains. @Lake2017 argue that human learning transfer abilities take advantage of important cognitive building blocks such as an abstract representation of concepts underlying tasks and compositionally structured causal models of the environment.

@@

Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, @ho1998iterated measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after training on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. @knez2000 test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to coordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.

@Juvina2013 made a similar distinction between what they deemed surface and deep similarities and find that both contribute to positive learning transfer. However, they show that surface similarity is not necessary for deep transfer and can either aid or block this type of transfer depending on whether it leads to congruent or incongruent actions in later games. In a series of experiments using economic signalling games Cooper & Kagel [-@cooper2003lessons; -@Cooper2008] found that participants who have learned to play according to a Nash Equilibrium in one game can transfer this to subsequent games, even though the actions consistent with Nash Equilibrium in later games are different.  They show that this transfer is driven by the emergence of sophisticated players who are able to represent the strategic implications of their actions and reason about the consequences of a different payoff structure on an opponent's actions.

Most studies fail to offer a formal explanation of transfer or a modelling framework that can explain the experimental observation of transfer between games and generalise it to extensive classes of games. A notable exception is the effort by @Haruvy2012 to specify a model of learning where players learn abstract rules that they can generalise and transfer across dissimilar games, rather than action choices that can only be used within the same game. Participants played ten games, presented in 4x4 normal form (matrix payoffs). Their results suggest that subjects do transfer learning over descriptively similar but strategically dissimilar games and that this learning transfer is significant. They also showed that players learn abstract aspects of the game that are then transferred to new settings. Their rule-learning model, based on @stahl1995players, was able to capture participants' dynamic behavior and shows that the propensity to select particular rules is perfectly transferred across games.

In exploring opponent modelling and learning transfer, most studies adopted two types of opponents in the experimental setting.  Either human participants were matched with other participants or they played against a computer algorithm. Computer opponents were generally programmed not to change their strategies over the course of the task, allowing better experimental control. One of the commonalities in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. The strategies implemented by the computer opponents had a style of play that was not "human-like" in the sense that humans are not very good at playing specific mixed strategies with any precision of at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that "play like humans", and one way of achieving that is to embed theses agents with human-like theory of mind abilities.  @simon1972theories explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally complex problems such as finding Nash equilibria. Instead, they will try to 'satisfice' by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in 'level-k' theory, first adopted by @stahl1995players. It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogeneous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment @arad201211. Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below, and so on.

@@

### Tasks
In the first experiment, the three games were rock-paper-scissors, fire-water-grass and the numbers game. A typical rock-paper-scissors game (hereafter RPS) is a 3x3 zero sum game, with a cyclical hierarchy between possible actions: rock blunts scissors, paper wraps rock, and scissors cut paper. If one player chooses an action which dominates their opponent's action, the player wins (receives a reward of 1) and the other player loses (receives a reward of -1). Otherwise it is a draw and both players receive a reward of 0. It has a unique MSNE consisting of randomly playing one of the three options each time.The second game is identical to Rock-Paper-Scissors in all but action labels. We call it Fire-Water-Grass (FWG): Fire burns grass, water extinguishes fire, and grass absorbs water. We are interested in exploring whether learning is transferred in a fundamentally similar game where the only difference is in the description of the choice actions.  Finally, the numbers game is a generalization of rock-paper-scissors. In the variant we use, 2 participants concurrently pick a number between 1 and 5. To win in this game, a participant needs to pick a number exactly 1 higher than the number chosen by the opponent. For example, if a participant thinks their opponent will pick 3, they ought to choose 4 to win the round. To make the strategies cyclical as in RPS, the game stipulates that the lowest number (1) beats the highest number (5), so if the participant thinks the opponent will play 5, then the winning choice is to pick 1. This game has a structure similar to RPS in which every action is dominated by exactly one choice. All other possible combination of choices that are not consecutive are considered ties. A win would add 1 point to the score of the player, while a loss deduces one point and a tie does not affect the score. Similar to RPS, the MSNE is to play each action with equal probability in a random way.

### Procedure
Participants played 3 games sequentially against the same computer opponent. The computer opponent either used a level-1 or level-2 strategy. Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. Each point worth £0.02. An example of the interface for the rock-paper-scissors game is provided in Figure\ \@ref(fig:feedback-rps-exp2).

@@

Looking at the aggregate scores (See Figure\ \@ref(fig:exp1-avg-scores-game) ), the RPS game had the lowest average score across participants (M = 0.289, SD = 0.348) followed by NUMBERS (M = 0.31, SD = 0.347) and finally the FWG game had the highest average score (M = 0.454, SD = 0.354).  Aggregate average scores for each game were significantly different from 0 (hypothesised value of random play) using parametric one sample t-tests (RPS: t(51) = 7.26, $p$ < 0.001 ;  FWG:  t(51) = 10.04 , $p$ < 0.001 ; NUMBERS: t(51) = 7.17, $p$ < 0.001).
To analyse within and between game learning, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) by 2 (block: first half, second half) repeated measures ANOVA with the first factor varying between participants. There was a main effect of Game ( F(2,100) = 8.54, $\eta^{2}$  = 0.05,  p < 0.001), showing that average scores varied significantly over the games. Post-hoc pairwise comparisons showed that performance in the FWG game was significantly higher than in the RPS game ( t(100) =3.78, p = 0.0008 ), and the performance in the NUMBERS game was significantly lower than FWG game ( t(100) = -3.32 , p = 0.0024). The score in RPS was not significantly different from the score in NUMBERS ( t(100) = 0.45 , p = 0.65). The main effect of Block ( F(1,50) = 22.51 , p < .001, $\eta^{2}$ = 0.03) shows that the average score in the first half of games (M = 0.29)  was significantly lower than in the second half  of the games played (M = 0.40), which translates to within-game learning. The main effect of Condition (F(1,50) = 5.44, p = .024, $\eta^{2}$ = 0.05) indicates that scores were higher against the level-1 player (M = 0.43) than against the level-2 player (M = 0.27). This indicates that it was harder for participants, on average, to exploit the strategy of the more sophisticated opponent (level-2) compared to that of the comparatively less sophisticated agent (level-1).

```{r exp1-3factor-plot, fig.cap = "Average scores by game, block and condition", fig.align = "center"}

@@

Finally, the analysis showed a significant block by game interaction ( F(2,100) = 6.92 , p = .002, $\eta^{2}$ = 0.02), indicating that within-game learning differed between the games. Indeed, second half scores in RPS are significantly higher than first-half scores (t(150) = 5.59, p <.0001), while there was no significant difference between block scores for the other two games. This is indicative of the significant within-game learning happening in the first game when players have no experience against the opponent, as opposed to much lower within game performance improvement in the latter games when participants have had some experience playing against the opponent and start with higher scores indicative of transfer.  There was also a three-way interaction between condition, game, and block ( F(2,100) = 3.88 , p = .023, $\eta^{2}$ = 0.01), which indicates, as seen in Figure\ \@ref(fig:exp1-3factor-plot) that within-game learning changes across games also depend on the sophistication of the opponent. For instance, there is more within game learning in the third game against level-2 opponents, since the initial scores are lower than against level-1 opponent. The explanation for this will become clearer when we discuss the factors moderating learning transfer in the next section.

@@

```{r exp1-score-by-opp, fig.cap = "Mean and CI of early scores by game and opponent", fig.align ='center'}

p2 <- ggerrorplot(exp1_dat2_6, x = "game.f", y = "early_score", group = 1, color="game.f", desc_stat = "mean_ci",palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","Numbers"), ylab = "Early rounds score", xlab = "Games", legend.title = "Games")

@@

```{r ANOVA early exp1, include = FALSE, warning=FALSE}

@@

#apa_lm1 <- apa_print(exp1_early_anova)

@@

# Second Experiment

We ran a second experiment with various differentiated features to improve the opportunity to measure learning transfer. Instead of making participants face either the level-1 or level-2 player throughout, we made them face both opponent sequentially. Because there were two distinct opponents, requiring potentially holding two opponent models in memory, we also made it easier to recall the results of past rounds by providing participants with the opportunity to see the history of the game since the beginning of each interaction. Figure\ \@ref(fig:feedback-rps-exp2) shows an example of showing interaction history in the RPS game. Finally, we changed the third game to a penalty shootout game, which has the same number of actions as the first two. If we see evidence for differential play against opponents, it would show participants adapting their strategies to the opponent they are facing, which is indicative of opponent modelling.

@@

```{r exp2-early-scores, fig.cap="Average early round scores and confidence intervals by game and opponent for experiment 2", fig.align="center"}
# plot scores per game 
p2 <- ggerrorplot(exp2_dat2_6, x = "game.f", y = "early_score", group = 1, color="game.f", desc_stat = "mean_ci",palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","SHOOT"), ylab = "Average Score rounds 2-6", xlab = "Games")

@@

$$ V_{t+1}(a) = V_{t}(a) + \alpha*( R_{t}  - V_{t}(a)) $$

@@

$$ Q^{new}(s_{t},a_{t}) = Q(s_{t},a_{t}) + \alpha*( R_{t}  + \gamma*\underset{a}{\max}Q(s_{t+1},a) - Q(s_{t},a_{t}) ) $$

@@

<!-- In experiment 2, we allow this richer model to have a memory of the interaction. In effect, we have a Q-learning model with a state space consisting of last round play, but in one instance, we fit a model (QLS) in which, when encountering an opponent, learning starts anew with no prior knowledge from previous encounters with the opponent used. In another instance, we fit a model in which what was learned about an opponent in the first interaction is transferred to the first round of the second interaction, so there is a within-game transfer of Q-values allowed (QLS-within). We fit both models to all participant's data in the second experiment where both opponents are faced.  -->

### EWA Models 
Next, we use a self-tuning Experience Weighted attraction model [@ho2004economics]. EWA models particularity is that they nest two seemingly different approaches, namely reinforcement learning and belief learning. Belief based models are based on the assumption that players keep track of the frequency of past plays and best respond to that. In contrast, reinforcement learning does not take into account beliefs about other players, but is such that an action followed by a positive reward is more likely to be repeated than an action followed by a negative reward. The self-tuning EWA model has been shown to perform better than both these nested models in multiple repeated games and has the advantage of having only one free parameter, the inverse temperature in the softmax choice function.

Let's define some notation in order to write the update rule of the self-tuning EWA model. For player i, there are $m_{i}$ strategies, denoted $s_{i}^{j}$ (i.e player i's strategy number j). Strategies actually played by i in period t, are denoted $s_{i}(t)$, while the opponent's strategy at time t is denoted $s_{-i}(t)$. After playing  $s_{i}^{j}$ at time $t$, player i pay-off is denoted $\pi_{i}(s_{i}^{j},s_{-i}(t))$, and the actual pay-off the player received is $\pi_{i}(t)$.

The EWA model is based on updating "Attractions" for each action over time. For instance, the attraction of strategy $j$ to player $i$ at time $t$ is written $A_{i}^{j}(t)$. Future action choice probabilities are based on these attractions using the softmax playing rule:

$$P_{i}^{j}(t+1) = \frac{e^{\lambda . A_{i}^{j}(t)}}{\sum_{k=1}^{m_{i}}e^{\lambda . A_{i}^{k}(t)}} $$

@@

$$ A_{i}^{j}(t) =  \frac{\phi . N(t-1) . A_{i}^{j}(t-1) + [ \delta + (1-\delta) . I(s_{i}^{j},s_{i}(t))] . \pi_{i}(s_{i}^{j},s_{-i}(t)) } {\phi . N(t-1) + 1} $$

Here, I(x,y) is the indicator function equal to 1 if $x = y$ and 0 otherwise. A simple way to think about this update rule is that attractions are multiplied by a parameter that represents experience ($N(t-1)$) which is itself decaying by a weight $\phi$. The result is added to either the pay-off received (when the indicator function is 0), or to $\delta$ times the foregone pay-off (when indicator function is 1).   We can see that setting $\delta = 0$ leads to reinforcement of past actions, while positive and high delta parameters make the update rule take into account foregone pay-offs, which is similar to weighted fictitious play [@cheung1994learning]. 
While the assumption in expanding the update rule above is that $\phi$ and $\delta$ are free parameters [@camerer1997experience], the self-tuning aspect of the model comes from the fact that these are now self-tuned using the formulas expanded in [@ho2004economics].

### ToM models

In this set of models, we assume that participants have a belief that the opponent is a level-k agent, with uniform probability of the level k, and use evidence of past play to update their beliefs in a Bayesian way about the true value of k. We use values of k  in ${0,1,2}$. These models also assume the opponent can deviate from these level-k strategies and play randomly with probability $\theta$, a parameter to be fit. We distinguish between multiple ToM models based on their ability to keep what was learned about the opponent in memory and hence facilitate transfer. In a No-Between-Transfer (NBT) model, players have no memory of what was learned about the opponent and start every new game assuming each level-k has equal probability. In the context of Experiment 2 where players face both opponents, this model assumes that participants transfer learning within the same game, from the first to the second interaction with the opponent, but are not able to transfer that learning to new games (So within but no between transfer)

Conversely, In a Between-Transfer model (BT), players are assumed to keep in memory what was learned about the type of opponent faced (vector of probabilities of level-k) and use that at the beginning of each new game. In the context of experiment 2, we still assume that if between transfer is present, then within game transfer is also present (from first to second interaction).

We fit the above two models to experiment 1 data. In experiment two, on top of these two models, we fit another model in which all stages of the game and all new games start with a fresh uniform probability of level-k opponent (NT), so no within or between opponent model learning transfer.

### Estimation and model comparison
In both experiments, all models were fit to each participant data, with optimal parameters being estimated using maximum likelihood. Using information criteria based Bayesian model comparison (BIC), the best fitting model for each participant was chosen and we compared the number of participants whose behavior was best explained by each model.

>>>>

---
title             : "Transfer of Learned Opponent Models in Zero Sum Games"
shorttitle        : "Upgrade Report"

@@

abstract: |
  Human learning transfer abilities take advantage of important cognitive building blocks such as an abstract representation of concepts underlying tasks and 
  causal models of the environment. One way to build abstract representations of the environment when the task involves interactions with others is to build 
  a model of the opponent that may inform what actions they are likely to take next. In this study, we propose to explore opponent modelling and its transfer with the use of
  computer agents possessing human-like theory of mind abilities with limited degrees of iterated reasoning. Through two differentiated experiments, we find that participants can
  deviate from Nash equilibrium play and learn to adapt to the opponent strategy and exploit it. Moreover, we showed that participants transfer their learning to new games and that
  the transfer is moderated by the level of  sophistication of the opponent. Computational modelling shows that it is likely that players start each game using a model-based
  learning strategy that facilitates generalisation and opponent model transfer, but then switch to behaviour that is consistent with a  model-free learning strategy in the latter
  stages of the interaction.

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
linkcolor         : "blue"
mask              : no
draft             : no
fig_caption       : yes

documentclass     : "apa6"
# classoption       : a4paper
classoption       : man
output            : papaja::apa6_pdf
header-includes:
  - \shorttitle{}
---

@@

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(plyr)
library(tidyr)
library(dplyr)
library(MASS)
library(ggpubr)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(magick)
library(depmixS4)

@@

```{r, include=FALSE} 
options(tinytex.verbose = TRUE)
#add_wordcount_filter(NULL, error = FALSE)
```

# Introduction
Being able to transfer previously acquired knowledge to a new domain is one of the hallmarks of human intelligence. Humans are naturally endowed with the ability to extract relevant features from a situation, identify the presence of these features in a novel setting and use previously acquired knowledge to adapt to previously unseen challenges using acquired knowledge. More formally, @perkins1992transfer define transfer of learning as the application of skills, knowledge, and/or attitudes that were learned in one situation to another learning situation. This typically human skill has so far eluded modern AI agents. Deep neural networks for instance can do very well on image recognition tasks and can even reach super-human performance levels on video and strategic board games. Yet they struggle to learn as fast or as efficiently as humans do, and more importantly they have a very limited ability to generalize and transfer knowledge to new domains. @Lake2017 argue that human learning transfer abilities take advantage of important cognitive building blocks such as an abstract representation of concepts underlying tasks and compositionally structured causal models of the environment.

@@

<!-- Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, @ho1998iterated measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after training on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. @knez2000 test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to coordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.   -->

<!-- @Juvina2013 made a similar distinction between what they deemed surface and deep similarities and find that both contribute to positive learning transfer. However, they show that surface similarity is not necessary for deep transfer and can either aid or block this type of transfer depending on whether it leads to congruent or incongruent actions in later games. In a series of experiments using economic signalling games Cooper & Kagel [-@cooper2003lessons; -@Cooper2008] found that participants who have learned to play according to a Nash Equilibrium in one game can transfer this to subsequent games, even though the actions consistent with Nash Equilibrium in later games are different.  They show that this transfer is driven by the emergence of sophisticated players who are able to represent the strategic implications of their actions and reason about the consequences of a different payoff structure on an opponent's actions. -->

In exploring opponent modelling and learning transfer, most studies adopted two types of opponents in the experimental setting.  Either human participants were matched with other participants or they played against a computer algorithm. Computer opponents were generally programmed not to change their strategies over the course of the task, allowing better experimental control. One of the commonalities in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. The strategies implemented by the computer opponents had a style of play that was not "human-like" in the sense that humans are not very good at playing specific mixed strategies with any precision, or at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that "play like humans", and one way of achieving that is to embed theses agents with human-like theory of mind abilities.  @simon1972theories explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally complex problems such as finding Nash equilibria. Instead, they will try to 'satisfice' by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in 'level-k' theory, first adopted by @stahl1995players. It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogeneous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment @arad201211. Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below, and so on.

@@

### Tasks
In the first experiment, the three games were rock-paper-scissors, fire-water-grass and the numbers game. A typical rock-paper-scissors game (hereafter RPS) is a 3x3 zero sum game, with a cyclical hierarchy between possible actions: rock blunts scissors, paper wraps rock, and scissors cut paper. If one player chooses an action which dominates their opponent's action, the player wins (receives a reward of 1) and the other player loses (receives a reward of -1). Otherwise it is a draw and both players receive a reward of 0. It has a unique MSNE consisting of randomly playing one of the three options each time. The second game is identical to Rock-Paper-Scissors in all but action labels. We call it Fire-Water-Grass (FWG): Fire burns grass, water extinguishes fire, and grass absorbs water. We are interested in exploring whether learning is transferred in a fundamentally similar game where the only difference is in the description of the choice actions.  Finally, the numbers game is a generalization of rock-paper-scissors. In the variant we use, 2 participants concurrently pick a number between 1 and 5. To win in this game, a participant needs to pick a number exactly 1 higher than the number chosen by the opponent. For example, if a participant thinks their opponent will pick 3, they ought to choose 4 to win the round. To make the strategies cyclical as in RPS, the game stipulates that the lowest number (1) beats the highest number (5), so if the participant thinks the opponent will play 5, then the winning choice is to pick 1. This game has a structure similar to RPS in which every action is dominated by exactly one choice. All other possible combination of choices that are not consecutive are considered ties. A win would add 1 point to the score of the player, while a loss deduces one point and a tie does not affect the score. Similar to RPS, the MSNE is to play each action with equal probability in a random way.

### Procedure
Participants played 3 games sequentially against the same computer opponent. The computer opponent either used a level-1 or level-2 strategy. Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. Each point is worth £0.02. An example of the interface for the rock-paper-scissors game is provided in Figure\ \@ref(fig:feedback-rps-exp2).

@@

Looking at the aggregate scores (See Figure\ \@ref(fig:exp1-avg-scores-game) ), the RPS game had the lowest average score across participants (M = 0.289, SD = 0.348) followed by NUMBERS (M = 0.31, SD = 0.347) and finally the FWG game had the highest average score (M = 0.454, SD = 0.354).  Aggregate average scores for each game were significantly different from 0 (hypothesised value of random play) using parametric one sample t-tests (RPS: t(51) = 7.26, $p$ < 0.001 ;  FWG:  t(51) = 10.04 , $p$ < 0.001 ; NUMBERS: t(51) = 7.17, $p$ < 0.001).
To analyse within and between game learning, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) by 2 (block: first half, second half) repeated measures ANOVA with the first factor varying between participants. There was a main effect of Game (F(2,100) = 8.54, $\eta^{2}$  = 0.05,  p < 0.001), showing that average scores varied significantly over the games. Post-hoc pairwise comparisons showed that performance in the FWG game was significantly higher than in the RPS game (t(100) =3.78, p = 0.0008), and the performance in the NUMBERS game was significantly lower than FWG game (t(100) = -3.32 , p = 0.0024). The score in RPS was not significantly different from the score in NUMBERS (t(100) = 0.45 , p = 0.65). The main effect of Block (F(1,50) = 22.51 , p < .001, $\eta^{2}$ = 0.03) shows that the average score in the first half of games (M = 0.29)  was significantly lower than in the second half  of the games played (M = 0.40), which translates to within-game learning. The main effect of Condition (F(1,50) = 5.44, p = .024, $\eta^{2}$ = 0.05) indicates that scores were higher against the level-1 player (M = 0.43) than against the level-2 player (M = 0.27). This indicates that it was harder for participants, on average, to exploit the strategy of the more sophisticated opponent (level-2) compared to that of the comparatively less sophisticated agent (level-1).

```{r exp1-3factor-plot, fig.cap = "Boxplots of scores by game, block and condition", fig.align = "center"}

@@

Finally, the analysis showed a significant block by game interaction ( F(2,100) = 6.92 , p = .002, $\eta^{2}$ = 0.02), indicating that within-game learning differed between the games. Indeed, second half scores in RPS are significantly higher than first-half scores (t(150) = 5.59, p <.0001), while there was no significant difference between block scores for the other two games. This is indicative of the significant within-game learning happening in the first game when players have no experience against the opponent, as opposed to much lower within game performance improvement in the latter games when participants have had some experience playing against the opponent and start with higher scores indicative of transfer.  There was also a three-way interaction between condition, game, and block ( F(2,100) = 3.88, p = .023, $\eta^{2}$ = 0.01), which indicates, as seen in Figure\ \@ref(fig:exp1-3factor-plot) that within-game learning changes across games also depend on the sophistication of the opponent. For instance, there is more within game learning in the third game against level-2 opponents, since the initial scores are lower than against level-1 opponent. The explanation for this will become clearer when we discuss the factors moderating learning transfer in the next section.

@@

(ref:figure4-caption) Mean and standard error of early scores by game and opponent
```{r exp1-score-by-opp, fig.cap = "(ref:figure4-caption)", fig.align ='center'}

p2 <- ggerrorplot(exp1_dat2_6, x = "game.f", y = "early_score", group = 1, color="game.f", desc_stat = "mean_se",palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","Numbers"), ylab = "Early rounds score", xlab = "Games", legend.title = "Games")

@@

```{r ANOVA early exp1, include = TRUE, warning=FALSE}

@@

exp1_early_anova
#apa_lm1 <- apa_print(exp1_early_anova)

@@

# Experiment 2

We ran a second experiment with various differentiated features to improve the opportunity to measure learning transfer. Instead of making participants face either the level-1 or level-2 player throughout, we made them face both opponents sequentially. Because there were two distinct opponents, requiring potentially holding two opponent models in memory, we also made it easier to recall the results of past rounds by providing participants with the opportunity to see the history of the game since the beginning of each interaction. Figure\ \@ref(fig:feedback-rps-exp2) shows an example of showing interaction history in the RPS game. Finally, we changed the third game to a penalty shootout game, which has the same number of actions as the first two. If we see evidence for differential play against opponents, it would show participants adapting their strategies to the opponent they are facing, which is indicative of opponent modelling.

@@

(ref:figure7-caption) Mean and standard error of early scores by game and opponent for experiment 2
```{r exp2-early-scores, fig.cap="(ref:figure7-caption)", fig.align="center"}
# plot scores per game 
p2 <- ggerrorplot(exp2_dat2_6, x = "game.f", y = "early_score", group = 1, color="game.f", desc_stat = "mean_se",palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","SHOOT"), ylab = "Early rounds score", xlab = "Games", legend.title = "Games")

@@

$$ V_{t+1}(a) = V_{t}(a) + \alpha \times ( R_{t}  - V_{t}(a)) $$

@@

$$ Q^{new}(s_{t},a_{t}) = Q(s_{t},a_{t}) + \alpha \times( R_{t}  + \gamma \times \underset{a}{\max}Q(s_{t+1},a) - Q(s_{t},a_{t}) ) $$

@@

## EWA Models 
Next, we use a self-tuning Experience Weighted attraction model [@ho2004economics]. EWA models particularity is that they nest two seemingly different approaches, namely reinforcement learning and belief learning. Belief based models are based on the assumption that players keep track of the frequency of past plays and best respond to that. In contrast, reinforcement learning does not take into account beliefs about other players, but is such that an action followed by a positive reward is more likely to be repeated than an action followed by a negative reward. The self-tuning EWA model has been shown to perform better than both these nested models in multiple repeated games and has the advantage of having only one free parameter, the inverse temperature in the softmax choice function. 
Let's define some notation in order to write the update rule of the self-tuning EWA model. For player i, there are $m_{i}$ strategies, denoted $s_{i}^{j}$ (i.e player i's strategy number j). Strategies actually played by i in period t, are denoted $s_{i}(t)$, while the opponent's strategy at time t is denoted $s_{-i}(t)$. After playing  $s_{i}^{j}$ at time $t$, player i pay-off is denoted $\pi_{i}(s_{i}^{j},s_{-i}(t))$, and the actual pay-off the player received is $\pi_{i}(t)$. 
The EWA model is based on updating "Attractions" for each action over time. For instance, the attraction of strategy $j$ to player $i$ at time $t$ is written $A_{i}^{j}(t)$. Future action choice probabilities are based on these attractions using the softmax playing rule:

$$P_{i}^{j}(t+1) = \frac{e^{\lambda  A_{i}^{j}(t)}}{\sum_{k=1}^{m_{i}}e^{\lambda A_{i}^{k}(t)}} $$

@@

$$ A_{i}^{j}(t) =  \frac{\phi \times N(t-1) \times A_{i}^{j}(t-1) + [ \delta + (1-\delta) \times I(s_{i}^{j},s_{i}(t))] \times \pi_{i}(s_{i}^{j},s_{-i}(t)) } {\phi \times N(t-1) + 1} $$

Here, $I(x,y)$ is the indicator function equal to 1 if $x = y$ and 0 otherwise. A simple way to think about this update rule is that attractions are multiplied by a parameter that represents experience ($N(t-1)$) which is itself decaying by a weight $\phi$. The result is added to either the pay-off received (when the indicator function is 0), or to $\delta$ times the foregone pay-off (when indicator function is 1).   We can see that setting $\delta = 0$ leads to reinforcement of past actions, while positive and high delta parameters make the update rule take into account foregone pay-offs, which is similar to weighted fictitious play [@cheung1994learning]. While the assumption in expanding the update rule above is that $\phi$ and $\delta$ are free parameters [@camerer1997experience], the self-tuning aspect of the model comes from the fact that these are now self-tuned using the formulas expanded in [@ho2004economics].

## BCH models

In this set of models, we assume that participants start with a prior belief that the opponent is a level-k agent, with uniform probability of the level $k \in \{ 0, 1, 2\}$, and use evidence of past play to update their beliefs in a Bayesian way about the true value of k. More formally, if we define $D_{t}$ as data available at time $t$ then the probability of the opponent being a level-k is updated using:

$$ P(level=k | D_{t})  \propto  P(D_{t}|level=k ) \times P(level=k)$$

These models also assume the opponent can deviate from these level-k strategies and play randomly with probability $\theta$, a parameter to be fit. We also assume that participants base their actions on their beliefs about the level of the opponent, so actions get chosen probabilistically using a softamx choice rule with probabilities of each opponent level as inputs. We distinguish between multiple BCH models based on their ability to keep what was learned about the opponent in memory and hence facilitate transfer. In a No-Between-Transfer (NBT) model, players have no memory of what was learned about the opponent and start every new game assuming each level-k has equal probability. In the context of Experiment 2 where players face both opponents, this model assumes that participants transfer learning within the same game, from the first to the second interaction with the opponent, but are not able to transfer that learning to new games (within but no between transfer).Conversely, In a Between-Transfer model (BT), players are assumed to keep in memory what was learned about the type of opponent faced (vector of probabilities of level-k) and use that at the beginning of each new game. In the context of experiment 2, we still assume that if between transfer is present, then within game transfer is also present (from first to second interaction).  In experiment two, on top of these two models, we fit another model in which all stages of the game and all new games start with a fresh uniform probability of level-k opponent (NT), so no within or between opponent model learning transfer.

## Estimation and model comparison
In both experiments, all models were fit to each participant data, with optimal parameters being estimated using maximum likelihood. Using information criteria based Bayesian model comparison (BIC), the best fitting model for each participant was chosen and we compared the number of participants whose behavior was best explained by each model.
